Earth's Impact History and the Timing of the Origin of Life

### 2023-04-19 ms.tex
discussion conclusion draft

\begin{abstract}
Big impacts on the early Earth would have created highly reducing atmospheres that generated molecules needed for the origin of life, such as nitriles. However, such impactors can be followed by impactors that were still sufficiently big to vaporize the ocean and destroy any incipient life. In this scenario, a post-impact reducing atmosphere needs to be followed by a lack of subsequent sterilizing impacts for life to persist. Using what we know about the impact history of early Earth and what is needed to generate post-impact highly reducing atmospheres, we show that  the median timing of biopoesis in this impact-driven scenario was $\sim 4.35$ Ga. However, uncertainties are large because impact bombardment is inherently stochastic, so biopoesis could have occurred in $\sim 0.5$ billion year window from 4.45 to 3.9 Ga within 95\% uncertainties of our model. Claims of a far narrower window for the origin of life are unsupported. In our optimistic scenario for life starting from post-impact reducing atmospheres, we find that biopoesis is possible in 92\% of stochastic impact realizations. This potentially bodes well for life on rocky planets elsewhere because these will have experienced an early episode of enhanced impact bombardment given how planets form.
\end{abstract}

@@

\citet{Benner_2020} argued that the most likely time for the origin of life in an RNA-world scenario would have been $4.36 \pm 0.1$ Ga in the wake of a $2 \times 10^{22}$ kg ($\sim 2100$ km) asteroid impact which they called Moneta. They suggested that iron delivered by Moneta's core would have reacted with impact-vaporized ocean water to generate a reducing Hadean atmosphere conducive to the photochemical generation of essential prebiotic molecules like HCN. This transient period could have been a window of opportunity for the origin of RNA and, ultimately, life.

@@

>>>>

\begin{abstract}
Big impacts on the early Earth would have created highly reducing atmospheres that generated molecules needed for the origin of life, such as nitriles. However, such impactors can be followed by impactors that were still sufficiently big to vaporize the ocean and destroy any incipient life. In this scenario, a post-impact reducing atmosphere needs to be followed by a lack of subsequent sterilizing impacts for life to persist. Using what we know about the impact history of early Earth and what is needed to generate post-impact highly reducing atmospheres, we show that  the median timing of biopoesis in this impact-driven scenario was $\sim 4.35$ Ga. However, uncertainties are large because impact bombardment is inherently stochastic, so biopoesis could have occurred in $\sim 0.5$ billion year window from 4.45 to 3.9 Ga within 95\% uncertainties of our model. Claims of a far narrower window for the origin of life are unsupported. In our optimistic scenario for life starting from post-impact reducing atmospheres, we find that biopoesis is possible in 92\% of stochastic impact realizations. This potentially bodes well for life on rocky planets elsewhere because they will have experienced an early episode of enhanced impact bombardment given how planets form.
\end{abstract}

@@

\citet{Benner_2020} argued that the most likely time for the origin of life in an RNA-world scenario would have been $4.36 \pm 0.1$ Ga in the wake of a $\sim 2 \times 10^{22}$ kg ($\sim 2100$ km) asteroid impact which they called Moneta. They suggested that iron delivered by Moneta's core would have reacted with impact-vaporized ocean water to generate a reducing Hadean atmosphere conducive to the photochemical generation of essential prebiotic molecules like HCN. This transient period could have been a window of opportunity for the origin of RNA and, ultimately, life.

@@

We use Monte-Carlo simulations of Earth's impact history to determine the most likely timing for the origin of life in an RNA-first processes. We make use of results of \citep{Wogan_2023}, who find that significant origin of life molecules, such as HCN, are produced in the Hadean atmosphere after a $> 4 \times 10^{20}$ kg impact in an optimistic case, and after a $> 5 \times 10^{21}$ kg impact in a pessimistic scenario. We consider both possibilities, and assume such impacts can cause life's origin so long as they are not followed by a smaller ocean-vaporizing collision that sterilizes the planet.

For either the optimistic or pessimistic cases, we find that the most likely timing for an origin of life impactor is $\sim 4.35$ Ga with a 95\% uncertainty spanning the entire Hadean eon (approximately 4.45 to 3.9 Ga). The large uncertainty is caused by the intrinsic stochastic nature of impacts. These results suggest that the \citet{Benner_2020} proposed timing of the origin of life, at $4.36 \pm 0.1$ Ga, is too narrow. Furthermore, the mass of the life starting impactor is most likely (69\% to 90\% probability) smaller than the Moneta impactor (Figure \ref{fig:time_and_mass}) proposed by \citet{Benner_2020}, because the size frequency distribution of asteroids prefers more frequent smaller impactors.

% Caveats
% While the modeling effort described above attempts to use the most current understanding of Earth's impact history and its effects,

Our simulations of Earth's impact history do not always result in an origin of life. There are some impact histories where a collision of sufficient mass to produce prebiotic molecules does not occur, or alternatively, does occur but primitive life is subsequently destroyed by an ocean-vaporizing impact that does not rekindle the biosphere. A life-starting impact occurs 92\% or 41\% of the time when we assume a $> 4 \times 10^{20}$ kg or $> 5 \times 10^{21}$ kg impact without later ocean-vaporization is required to start life, respectively.

If life began in an RNA process after an impact, then this work suggests that origin of life on Earth was not an anomalous event and was indeed strongly favored (92\% chance) under optimistic assumptions. Given that rocky exoplanets accreted from impacts, our work supports and optimistic outlook for life on exoplanets and future searches for biosignatures.
